---
title: "Introduction to hansard"
author: "Evan Odell"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{hansard}
  %\VignetteEngine{knitr::rmarkdown}
    \usepackage[utf8]{inputenc}  
---

`hansard` is an R package to pull data from the UK parliament through the <http://www.data.parliament.uk/> API. It emphasises simplicity and ease of use, so that users unfamiliar with APIs can easily retrieve large volumes of high quality data. Each function accepts a single argument at a time, and functions that require additional information to retrieve the data you requested will ask for it after you execute the function. Functions retrieve data in json format and convert it to a data frame. The `hansard_generic` function supports the building of API requests for XML, csv or HTML formats if required. Note that the API is rate limited to returning 5500 rows per request in some circumstances.

##Installing hansard

###From CRAN

```
install.packages("hansard")`
```

###From GitHub (Development Version)

```
install.packages("devtools")
devtools::install_github("EvanOdell/hansard")
```

###Load hansard

```
library(hansard)
```

##Using hansard

`hansard` contains functions for calling data for the UK Parliament API. Most of the functions are designed to call data from a specific <http://www.data.parliament.uk/> API, but several functions have special or experimental features:

* `research_briefings_lists` An experimental feature that creates lists of possible parameters, and allows the user to call to the API based on the content of the list. Currently only available with the `research_briefings_lists` function.

* `hansard_generic` A basic wrapper function for any of the available APIs. Users must specify the exact path they want the function to call from, and must also specify JSON formatting.

* `members_search` A search function for members of the House of Commons and House of Lords.

### The `members_search` function

Looking up information on an individual MP or Lord through the Parliamentary API requires knowing their parliamentary ID number. This can be hard to find on the web, but luckily you can look it up through the API. We want information on the voting record of the Labour MP for Hackney North and Stoke Newington Diane Abbott, but we don't know her ID number, so we search for her:

```{r echo=TRUE}
library(hansard)
members_search("abbot")

```


The same function, without the use of the tidy parameter, illustrating the differences in the variable names and the presentation of information in the first column:


```{r, echo=TRUE}
library(hansard)
members_search("abbot", tidy = FALSE)
```


The same function as above, but with `tidy_style = "period.case"`, so it returns a different 

```{r, echo=TRUE}
library(hansard)
members_search("abbot", tidy = TRUE, tidy_style = "period.case")
```


The search function is not case sensitive, and searchs in the names and constituencies of all MPs and Lords. So even though we spelled her surname incorrectly, we can still find her. This API provides limited biographical details, to retrieve more detailed biographical information, use the related [`mnis`](https://cran.r-project.org/package=mnis) package to retrives data from the Members' Names Information Service.


### The `commons_divisions` and `mp_vote_record` functions

The `commons_divisions` function returns divisions in the House of Commons, including the result of votes and information on what we being voted on. `mp_vote_record` returns a data frame the voting record of a given MP on each division they voted in. The example below returns all Commons Divisions where Diane Abbott voted aye.

```{r eval = FALSE}
x <- mp_vote_record(172, "aye")
#> Retrieving page 1 of 4
#> Retrieving page 2 of 4
#> Retrieving page 3 of 4
#> Retrieving page 4 of 4
#> head(x)
#> # A tibble: 6 × 5
#>                                         about                                                   title               uin
#>                                       <chr>                                                   <chr>             <chr>
#> 1 http://data.parliament.uk/resources/714865 Pension Schemes Bill [Lords]: Report Stage New Clause 2 CD:2017-03-29:260
#> 2 http://data.parliament.uk/resources/714866 Pension Schemes Bill [Lords]: Report Stage New Clause 1 CD:2017-03-29:259
#> 3 http://data.parliament.uk/resources/714868  Pension Schemes Bill [Lords]: Report Stage Amendment 1 CD:2017-03-29:261
#> 4 http://data.parliament.uk/resources/713962    Bus Services Bill [Lords]: Report Stage New Clause 1 CD:2017-03-27:255
#> 5 http://data.parliament.uk/resources/713963    Bus Services Bill [Lords]: Report Stage New Clause 2 CD:2017-03-27:256
#> 6 http://data.parliament.uk/resources/714005     Bus Services Bill [Lords]: Report Stage Amendment 1 CD:2017-03-27:257
#> # ... with 2 more variables: date_value <date>, date_datatype <fctr>
> 

```

The results of votes in the House of Lords can be retrieved with the `lords_divisions` function. The voting record of individual Lords can  be retrieved using the `lords_vote_record` functions. 


### The `research_briefings` function

The `research_briefings` function includes the experimental feature of requesting data using lists created using the `research_briefings_lists` functions:

```{r eval = FALSE}
#> research_topics_list <- research_topics_list()
#> 
#> research_subtopics_list <- research_subtopics_list()
#> 
#> research_types_list <- research_types_list()
#>
#> research_topics_list[[7]]
#> [1] "Defence"
#>
#> research_subtopics_list[[7]][10]
#> [1] "Falkland Islands"
#>
#> research_types_list[[1]]
#> [1] "Lords Library notes"

```

In this case I have given them the same name as their function, but you can assign any name you wish to them.

Having created the lists, they can be used to specify which topics and subtopics to call, although strings can also be used.

```{r eval = FALSE}
#> x <- research_briefings(topic = research_topics_list[[7]])
#>
#> x <- research_briefings(topic = research_topics_list[[7]], subtopic=research_subtopics_list[[7]][10])
#>
#> x <- research_briefings(topic = "Defence")

```

If a specific subtopic is called, but the topic is not specified, the function will still return all data within that specific subtopic. Note that this is slower than specifying the topic and subtopic.

```{r eval = FALSE}
#> x <- research_briefings(subtopic = research_subtopics_list[[7]][10])
#> 
#> x <- research_briefings(subtopic = "Falkland Islands")
#>
#> system.time(without_topic <- research_briefings(subtopic = research_subtopics_list[[7]][10]))
#> Retrieving page 1 of 1
#>   user  system elapsed 
#>   1.12    2.59    4.71 
#>
#> system.time(with_topic <- research_briefings(topic = research_topics_list[[7]], subtopic=research_subtopics_list[[7]][10]))
#> Retrieving page 1 of 1
#>    user  system elapsed 
#>    0.47    1.31    1.89 
#>
#> all.equal(with_topic, without_topic)
#> [1] TRUE
```


If a specified subtopic is not a subtopic of the specified topic, the function will not return any data.

### The `hansard_generic` function

The `hansard_generic` function allows you to put in your own paths to the API. Information on all the paths available in the API can be found on the [DDP Explorer website](http://explore.data.parliament.uk/).

```{r, eval=FALSE}
x <- hansard_generic("commonsansweredquestions.json")
```

Note that the API defaults to returning 10 items per page, but allows up to 500 items per page.
