---
title: "Introduction to Hansard"
author: "Evan Odell"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Hansard Introduction}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

`hansard` is an R package to pull data from the UK parliament through the <http://www.data.parliament.uk/> API. It emphasises simplicity and ease of use, so that users unfamiliar with APIs can easily retrieve large volumes of high quality data. Each function accepts a single argument at a time, and functions that require additional information to retrieve the data you requested will ask for it after you execute the function. Functions retrieve data in json format and convert it to a data frame. The `hansard_generic` function supports the building of API requests for XML, csv or HTML formats if required.

##Installing hansard

###From CRAN

install.packages("hansard")

###From GitHub (Development Version)

install.packages("devtools")
devtools::install_github("EvanOdell/hansard")

###Load hansard

library(hansard)

##Using hansard

`hansard` contains are 42 functions for calling data. Most of the functions are 

`all_answered_questions`
`bill_stage_types`
`bills`
`committee_reports`
`commons_answered_questions`
`commons_committee_inquiries`
`commons_division_date`
`commons_divisions`
`commons_interests`
`commons_members`
`commons_oral_question_times`
`commons_oral_questions`
`commons_terms`
`commons_written_questions`
`constituencies`
`early_day_motions`
`election_results`
`elections`
`epetition`
`hansard_basic`
`hansard_generic`
`lord_vote_record`
`lords_amendments`
`lords_attendance`
`lords_divisions`
`lords_interests`
`lords_members`
`lords_written_questions`
`members`
`members_search`
`mp_edms`
`mp_questions`
`mp_vote_record`
`papers_laid`
`publication_logs`
`research_briefings`
`research_topics`
`research_types`
`sessions_info`
`tv_channels`
`tv_clips`
`tv_programmes`


###Using members_search
Looking up information on an individual MP or Lord through the Parliamentary API requires knowing their parliamentary ID number. This can be hard to find on the web, but luckily you can look it up through the API. We want information on the voting record of the Labour MP for Hackney North and Stoke Newington Diane Abbott, but we don't know her ID number, so we search for her:

```{r eval=FALSE}
library(hansard)
members_search("abbot")
#>   MP.ID additionalName                      constituency constituencyCode
#> 1   172          Julie Hackney North and Stoke Newington           146966
#> 2  1651      Granville                              <NA>               NA
#> 3  4249           <NA>                      Newton Abbot           147092
#> 4  3827         Edmond                              <NA>               NA
#>   familyName                       fullName gender  givenName
#> 1     Abbott                Ms Diane Abbott Female      Diane
#> 2    Hodgson Lord Hodgson of Astley Abbotts   Male      Robin
#> 3     Morris              Anne Marie Morris Female Anne Marie
#> 4  Neuberger   Lord Neuberger of Abbotsbury   Male      David
#>                            homePage
#> 1     http://www.dianeabbott.org.uk
#> 2                              <NA>
#> 3 http://www.annemariemorris.co.uk/
#> 4       http://www.judiciary.gov.uk
#>                                                      label        party
#> 1                Biography information for Ms Diane Abbott       Labour
#> 2 Biography information for Lord Hodgson of Astley Abbotts         <NA>
#> 3              Biography information for Anne Marie Morris Conservative
#> 4   Biography information for Lord Neuberger of Abbotsbury         <NA>
#>                             twitter
#> 1 https://twitter.com/HackneyAbbott
#> 2                              <NA>
#> 3    https://twitter.com/AMMorrisMP
#> 4                              <NA>

```
The search function is not case sensitive, and searchs in the names and constituencies of all MPs and Lords. So even though we spelled her surname incorrectly, we can still find her.


###Using commons_divisions and mp_vote_record

The `commons_divisions` function returns divisions in the House of Commons and the divisions that a member votes no or aye in. The example below returns all Commons Divisions where Diane Abbott voted aye.

```{r eval = FALSE}
x <- commons_divisions("aye")
#> Enter Member ID: 172
#> Retrieving page 1 of 4
#> Retrieving page 2 of 4
#> Retrieving page 3 of 4
#> Retrieving page 4 of 4
#> head(x)
#>  _about                                                              title
#> 1 http://data.parliament.uk/resources/626911                            Opposition Motion: Community Pharmacies
#> 2 http://data.parliament.uk/resources/626953                           Opposition Motion: Police Officer Safety
#> 3 http://data.parliament.uk/resources/621252                                           Opposition Motion: Yemen
#> 4 http://data.parliament.uk/resources/607490 Closure Motion: Sexual Offences (Pardons Etc.) Bill second reading
#> 5 http://data.parliament.uk/resources/582798      Opposition Motion: NHS sustainabiliy and transformation plans
#> 6 http://data.parliament.uk/resources/576587                                Finance Bill: Report Stage Amdt 174
#>                 uin date._value date._datatype
#> 1 CD:2016-11-02:142  2016-11-02       dateTime
#> 2 CD:2016-11-02:143  2016-11-02       dateTime
#> 3 CD:2016-10-26:139  2016-10-26       dateTime
#> 4 CD:2016-10-21:138  2016-10-21       dateTime
#> 5 CD:2016-09-14:134  2016-09-14       dateTime
#> 6 CD:2016-09-06:125  2016-09-06       dateTime

```

The voting record of MPs and Lords can also be retrieved using the `mp_vote_record` and `lords_vote_record` functions, respectively. This function has several advantages over the `commons_divisions` and `lords_divisions` functions. All the required parameters for the API call are included in the function parameters, rather than using console input, and they can return all divisions that an MP or Lord voted in.

```{r eval = FALSE}
x <- mp_vote_record(172, "all")
#> Retrieving aye votes:
#> Connecting to API
#> Retrieving page 1 of 4
#> Retrieving page 2 of 4
#> Retrieving page 3 of 4
#> Retrieving page 4 of 4
#> Retrieving no votes:
#> Connecting to API
#> Retrieving page 1 of 3
#> Retrieving page 2 of 3
#> Retrieving page 3 of 3
#> head(x)
#>                                       _about                                                                       title
#> 1 http://data.parliament.uk/resources/653644 Opposition Motion: The Government's plan for Brexit (Prime Minister's Amdt)
#> 2 http://data.parliament.uk/resources/653645       Opposition motion: The Government's plan for Brexit motion as amended
#> 3 http://data.parliament.uk/resources/646440                             Digital Economy Bill: Report Stage New Clause 8
#> 4 http://data.parliament.uk/resources/641464                            Opposition Motion: Education and Social Mobility
#> 5 http://data.parliament.uk/resources/641522                          Opposition Motion: National Health Service Funding
#> 6 http://data.parliament.uk/resources/640531               Higher Education and Research Bill: Report Stage New Clause 2
#>                 uin date._value date._datatype vote
#> 1 CD:2016-12-07:169  2016-12-07       dateTime  aye
#> 2 CD:2016-12-07:170  2016-12-07       dateTime  aye
#> 3 CD:2016-11-28:165  2016-11-28       dateTime  aye
#> 4 CD:2016-11-22:161  2016-11-22       dateTime  aye
#> 5 CD:2016-11-22:162  2016-11-22       dateTime  aye
#> 6 CD:2016-11-21:154  2016-11-21       dateTime  aye

```


###Using the hansard_basic function

The `hansard_basic` function allows you to put in your own paths to the API. Information on all the paths available in the API can be found on the [DDP Explorer website](http://explore.data.parliament.uk/).

```{r, eval=FALSE}
x <- hansard_generic("commonsansweredquestions.json")
```

Note that the API defaults to returning 10 items per page, but allows up to 500 items per page. The other `hansard` functions include `"_pageSize=500"` as a suffix in the path to reduce the number of page calls required.
